SDR enhancement by clouds

### 2024-04-06 article/article.Rmd
Commit 2024-04-06 17:15

We approximated the expected clear sky GHI by using a model.  Because of the lack of
observational data for the whole period, we used some long term climatological data
for the main factors responsible for the attenuation of the broadband downward solar
radiation in the atmosphere. Which are mainly, the aerosols and the water vapors.
Fortunately, our site participates in the Aerosol Robotic Network (AERONET)
[@Giles2019; @Buis1998], as we operate a Cimel photometer since 2003, collocated with
the CM-21 pyranometer.  The mean monthly aerosol optical depth (AOD) on different
wavelengths is provided by AERONET, along with the equivalent water column height in
the atmosphere.  The simulation of the clear sky radiation was performed by the well
established radiation transfer model Libradtran [@Emde2016], a similar
approach, was also used by @Vamvakas2020 for creating a clear sky reference.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated for combinations of AOD at
$500\,\text{nm}$ ($t_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w_h$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

@@

To select the exact values of these thresholds factors (Equation \ref{eq:CE4}), we
implemented an empirical method, by manual inspection of the CE identification, on
specially selected days from the whole dataset.  We used seven sets of selected days,
with characteristics relevant to the efficiency of the identification threshold.
These sets were random groups of about 20 to 30 days with the following
characteristics:
(a) the largest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) without clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) without clouds and without EC events,
(e) with at least $60\%$ of the day length without clouds and some EC events,
(h) random days and
(i) some manual selected days that were included during the manual inspection.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
$t_{\text{cs}} = t_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{mCSi}$).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds.  To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.
-->
\begin{equation}
\text{CE} : \text{GHI}_\text{i} > \text{GHI}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{GHI}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.

@@

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and satisfy the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times \text{TSI}_\text{iTOA}
\label{eq:ECE}
\end{equation}

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the 'maximum likelihood' fitting method [@Gardner1980;
@Jones1980] by implementing the function 'arima' from the library 'stats' of the R
programming language [@RCT2023]. The trends are reported together with the $2\sigma$
error.  We observe an increase of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,W/m^2$
on the mean OIR.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  The annual
mean OIR is
$`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2$
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

A better indicator of changes on the characteristic of CE would be the number of CE
occurrences and the total energy of the CE over irradiance. The annual number of CE
occurrences, shows a steady increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,/y$
(Figure\nobreakspace{}\ref{fig:P-energy-N}).
We have to note that the energy related to the CE events can not be directly linked
with the total energy balance on the atmosphere. The net sun radiation of the region
is not increased, but rather redistributed through the CE. Although, the
instantaneous values, can exceed the equivalent clear sky irradiance to a
considerable level.

>>>>

We approximated the expected clear sky GHI by using a radiation transfer model.  The
simulations was performed by the well established model Libradtran [@Emde2016], a
similar approach, was also used by @Vamvakas2020 for creating a clear sky reference.
Because of the lack of observational data for the whole period, we used some long
term climatological data for the main factors responsible for the attenuation of the
broadband downward solar radiation in the atmosphere. Which are mainly, the aerosols
and the water vapors.  Fortunately, our site participates in the Aerosol Robotic
Network (AERONET) [@Giles2019; @Buis1998], as we operate a Cimel photometer since
2003, collocated with the CM-21 pyranometer.  The mean monthly aerosol optical depth
(AOD) on different wavelengths is provided by AERONET, along with the equivalent
water column height in the atmosphere.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated for combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w_h$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

@@

To select the exact values of these thresholds factors (Equation \ref{eq:CE4}), we
implemented an empirical method, by manual inspection of the CE identification, on
specially selected days from the whole dataset.  We used seven sets of selected days,
with characteristics relevant to the efficiency of the identification threshold.
These sets were random groups of about 20 to 30 days with the following
characteristics:
(a) the largest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) without clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) without clouds and without EC events,
(e) with at least $60\,\%$ of the day length without clouds and some EC events,
(h) random days and
(i) some manual selected days that were included during the manual inspection.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{mCSi}$).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds.  To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.
-->
\begin{equation}
\text{CE} : \text{GHI}_\text{i} > \text{GHI}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{GHI}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.

@@

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and satisfy the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, $i$ each of the one-minute observation

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the 'maximum likelihood' fitting method [@Gardner1980;
@Jones1980] by implementing the function 'arima' from the library 'stats' of the R
programming language [@RCT2023]. The trends are reported together with the $2\sigma$
error.  We observe an increase of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,W/m^2/y$
on the mean OIR.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  The annual
mean OIR is
$`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2/y$
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

A better indicator of changes on the characteristic of CE would be the number of CE
occurrences and the total energy of the CE over irradiance. The annual number of CE
occurrences, shows a steady increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/y$
(Figure\nobreakspace{}\ref{fig:P-energy-N}).
We have to note that the energy related to the CE events can not be directly linked
with the total energy balance on the atmosphere. The net sun radiation of the region
is not increased, but rather redistributed through the CE. Although, the
instantaneous values, can exceed the equivalent clear sky irradiance to a
considerable level.
